Vax Track Quebec
================

Tracking how quickly Quebec is moving its vaccine supply into human beings. All
data are taken from [covid19tracker.ca](https://covid19tracker.ca). This 
analysis is at the dose level and does not distinguish between first and second
doses, or vaccine types.

Some provinces have dates where they appear to have administered more doses
than were distributed to them. These doses are not included in wait-time counts.

## Setup

<details><summary>Click here to see the setup code</summary>

Load packages and import data:

```{r Setup, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
setwd('~/MyProjects/COVID19/VaxTrack/')
library(jsonlite)
library(tidyverse)
theme_set(theme_bw())
library(lubridate)
url = 'https://api.covid19tracker.ca/reports/province/qc?fill_dates=true'
upd = fromJSON(url)
dat_can = upd$data
# head(dat_can)
# tail(dat_can)
# summary(dat_can)
dat_can$date = ymd(dat_can$date)
source('functions.R')
# summary(dat_can)

dat_vax = (dat_can
		   %>% filter(!is.na(total_vaccines_distributed))
		   %>% select(date,change_vaccinations,
		   		   change_vaccines_distributed,total_vaccinations,
		   		   total_vaccines_distributed))
# dim(dat_vax)
# head(dat_vax)
```

[functions.R](./functions.R)
</details>

## Vaccine Gap

<details><summary>code</summary>
```{r}
dist_tot = dat_vax[nrow(dat_vax),'total_vaccines_distributed']
admn_tot = dat_vax[nrow(dat_vax),'total_vaccinations']
msg = paste('So far Quebec has received ', 
			format(dist_tot, nsmall = 1, big.mark = ','), 
			' vaccine doses, and ',
			format(admn_tot, nsmall = 1, big.mark = ','), 
			' of those have been administered.', sep = '')
msg1 = paste('This is a cumulative administration rate of ', 
			100*round(admn_tot/dist_tot, 4), '%', sep = '')
```
</details>

```{r, echo = FALSE}
msg
msg1
```

<details><summary>code</summary>
```{r}
# head(dat_vax)
dat_tot_long = (dat_vax
				%>% select(-change_vaccinations, 
						   -change_vaccines_distributed)
				%>% gather(Variable, Count, -date)
				%>% mutate(Variable = factor(Variable, 
											 levels = c('total_vaccines_distributed',
											 		   'total_vaccinations'))))
# head(dat_tot_long)
abs_adm = ggplot(dat_tot_long, aes(date, Count, fill = Variable)) +
	geom_area(position = 'identity') +
	scale_fill_manual(values = c('sienna4', 'sienna1'), 
					  labels = c('doses distributed','doses administered'),
					  name = NULL) +
	xlab('Date') +
	ggtitle('Total cumulative doses distributed\nand administered over time')
dat_vax = (dat_vax
		   %>% mutate(prop_dist = 
		   		   	total_vaccinations/total_vaccines_distributed,
		   		   prop_7 = roll_avg(prop_dist, 7)))
prop_adm = ggplot(dat_vax, aes(date, prop_dist)) +
	geom_line() +
	ylim(0,1) +
	xlab('Date') +
	ylab('Proportion Administered') +
	ggtitle('Instantaneous proportion of delivered doses\nthat have been administered')

prop_adm_7 = ggplot(dat_vax, aes(date, prop_7)) +
	geom_line() +
	ylim(0, 1) +
	xlab('Date') +
	ylab('Proportion Administered') +
	ggtitle('Rolling 7-day average of proportion of delivered doses\nthat have been administered')
```
</details>

```{r, echo = FALSE, warning=FALSE, fig.width=8.65}
abs_adm
```
```{r, echo = FALSE, warning=FALSE}
prop_adm
prop_adm_7
```


## Wait Times

<details><summary>code</summary>
```{r}
wt_df = get_wait(dat_vax)
wait = get_avg(wt_df)
msg = paste('Vaccine doses have waited an average of', round(wait,1), 'days between being distributed and being administered.')
```
</details>

```{r, echo = FALSE}
msg
```

<details><summary>code</summary>
```{r}
wait_df = data.frame(Wait = NA)

for(i in 2:nrow(dat_vax)){
	wait_df = rbind(wait_df, 
					data.frame(Wait = get_avg(wt_df, 1, i)))
}
wait_df = data.frame(Wait = wait_df[-1,])
head(wait_df)
wait_df$Date = dat_vax[-1,'date']

wait_plt = ggplot(wait_df, aes(Date, Wait)) +
	geom_line() +
	ylab('Wait Time (Days)') +
	ggtitle('Cumulative average time it takes a vaccine dose\nto go from distribution to administration\nin Quebec')
```
</details>

```{r, echo = FALSE}
wait_plt
```

<details><summary>code</summary>
```{r}
counted = sum(wt_df$Doses)
for (i in 1:nrow(dat_vax)){
	dist = dat_vax[i, 'total_vaccines_distributed']
	if (counted < dist){
		break
	}
}
n = nrow(wt_df)
leftover = dist - counted
pos = i
timediff = today() - dat_vax[pos, 'date']

avg_wt = data.frame(Time = timediff, Doses = leftover)
# avg_wt

for (i in (pos+1):nrow(dat_vax)){
	i = i + 1
	i
	dat_vax[i,]
	timediff = today() - dat_vax[i, 'date']
	timediff
	avg_wt = rbind(avg_wt,
				   data.frame(Time = timediff,
				   		   Doses = dat_vax[i, 'change_vaccines_distributed']))
	avg_wt
}

avg_wt = na.omit(avg_wt)

new_wt = sum(avg_wt$Time * avg_wt$Doses)/sum(avg_wt$Doses)

msg = paste('Current unadministered doses have been waiting an average of',
			round(new_wt,1), 'days.')
```
</details>

```{r, echo = FALSE}
msg
```

---

<details><summary>code</summary>
```{r}
msg = paste('This website was last updated', today())
msg1 = paste('The most recent data are from', max(dat_vax$date))
```
</details>

```{r, echo = FALSE}
msg
msg1
```

This website is normally automatically updated at 11pm EDT daily. Exceptions 
may occur.
